---
title: "Deliverable2"
author: "Sarah Hyland"
format: docx
editor: visual
---

# Install packages

```{r}
#install.packages("tidyverse")
#install.packages("readxl")
#install.packages("GGally")
```

# Load packages

```{r}
library(tidyverse)
library(readxl)
library(GGally)
```

# Get list of files in the folder

```{r}
list.files("C:/Users/Lenovo/OneDrive/Desktop/GIT/deliverable2/data")
```

# Sheets in the Excel workbook

```{r}
excel_sheets("C:/Users/Lenovo/OneDrive/Desktop/GIT/deliverable2/data/rockydata.xlsx")
```

# Read in data from sheet 'Sites'

```{r}
sites <- read_xlsx("C:/Users/Lenovo/OneDrive/Desktop/GIT/deliverable2/data/rockydata.xlsx", sheet = "Sites")
```

## Summarise Sites

```{r}
sites
```

# Read in data from sheet 'Species'

```{r}
species <- read_xlsx("C:/Users/Lenovo/OneDrive/Desktop/GIT/deliverable2/data/rockydata.xlsx", sheet = "Species")
```

## Summarise Species

```{r}
species
```

# Getting thing tidy

## Putting site data into long format

```{r}
sites %>%
pivot_longer(cols = c("PercentRock", "waterPH", "SurfaceTemp"), names_to = "variable", values_to = "value")
```

## Putting species data into wide format

```{r}
species %>%
select(Site, Point, WorkingName) %>%
mutate(Presence = 1) %>% # adds a column called "Presence" filled with "1"
  pivot_wider(names_from = WorkingName, values_from = Presence, values_fill = 0)
```

## Use unite to combine the site and point columns into a new column called SitePoint

```{r}
read_xlsx("C:/Users/Lenovo/OneDrive/Desktop/GIT/deliverable2/data/rockydata.xlsx", sheet = "Species") %>% 
  unite("SitePoint", Site:Point, sep = "_") %>%
  select(SitePoint, WorkingName) %>%
  mutate(Presence = 1) %>% # adds a column called "Presence" filled with "1"
  pivot_wider(names_from = WorkingName, values_from = Presence, values_fill = 0)
```

## Calculate number of species in each site

```{r}
sr <- read_xlsx("C:/Users/Lenovo/OneDrive/Desktop/GIT/deliverable2/data/rockydata.xlsx", sheet = "Species") %>% 
  unite("SitePoint", Site:Point, sep = "_") %>%
  select(SitePoint, WorkingName) %>% 
  group_by(SitePoint) %>%
  summarize(`Species Number` = n())

sr
```

## Read in site data, select columns and join sr

```{r}
read_xlsx("C:/Users/Lenovo/OneDrive/Desktop/GIT/deliverable2/data/rockydata.xlsx", sheet = "Sites") %>%
  select(SitePoint, PercentRock, waterPH) %>%
  left_join(sr, by = "SitePoint")
```

# Plot comparisons

```{r}
read_xlsx("C:/Users/Lenovo/OneDrive/Desktop/GIT/deliverable2/data/rockydata.xlsx", sheet = "Sites") %>%
  select(SitePoint, PercentRock, waterPH) %>%
  left_join(sr, by = "SitePoint") %>%
  GGally::ggpairs(columns = 2:ncol(.))
```
